---
title: "Gov 50 Final Project"
author: "Violet Barron"
description: "My final project"
output:
  distill::distill_article:
    self_contained: false
---



## Project thoughts

I am interested in exploring data related to the vast disparities between various Asian American and Native Hawaiian/Pacific Islander (AANHPI) ethnic subgroups in the United States. I plan to use publicly available U.S. census data (found at data.census.gov) to explore the varying levels of educational attainment found across these AANHPI ethnic groups and then measure these observations against the national averages for educational attainment. Through this project, I hope to disprove the reductionist view of the AANHPI community as an exclusively  well-educated monolith.

## Data source

How do different Asian American and Native Hawaiian/Pacific Islander (AANHPI) ethnic groups fare in America when their experiences are measured against national averages? In this study, I plan to explore the discrepancy in experiences — measured through levels of educational attainment — between various AANHPI ethnic groups and compare this discrepancy with the experience of the average American citizen. I hypothesize that there will not be a consensus as to whether or not AANHPI on the whole are more formally educated than the average American citizen, and rather certain AANHPI ethnicities, such as Chinese Americans, will be more formally educated than the average American while other AANHPI ethnicities, such as Hmong Americans, will be less formally educated than the average American citizen. My sample is comprised of U.S. census data from 2010-2020. My unit of analysis is a specific AANHPI ethnic group and for this project, I will only consider Chinese and Hmong Americans as while these two ethnicities are often lumped into one, I hypothesize that these two groups will have perhaps the most dissonant experiences when it comes to educational attainment. The explanatory variable of interest is whether each of these ethnic groups fare better or worse than the national average for educational attainment level. The variable is measured from publicly available U.S. data (found at data.census.gov). If I observe that across ethnic groups, the aggregate of the AANHPI population fare about equally to the average American, this would provide support for my hypothesis, as the groups that fare better would "cancel out" those that fare worse, leading to an overall educational attainment level that is about average. If, on the other hand, I observe that across ethnic groups, the aggregate of the AANHPI population fare either overwhelmingly better or worse than the average American citizen, this would provide evidence against my hypothesis. When I run my regression of the aggregate AANHPI experience against the national average, a positive, relatively insignificant coefficient would indicate support for my hypothesis.

## Milestone 3: Example use of disagreggated census data

```{r include = FALSE}
library(tidyverse)
acs2022 <- read_csv("data/acs2022.csv")

acs2022select <- acs2022 |> select(-"United States!!Asian Indian alone!!Margin of Error",
                                   -"United States!!Bangladeshi alone!!Margin of Error",
                                   -"United States!!Cambodian alone!!Margin of Error",
                                   -"United States!!Chinese alone!!Margin of Error",
                                   -"United States!!Chinese, except Taiwanese alone!!Margin of Error",
                                   -"United States!!Taiwanese alone!!Margin of Error",
                                   -"United States!!Filipino alone!!Margin of Error",
                                   -"United States!!Hmong alone!!Margin of Error",
                                   -"United States!!Indonesian alone!!Margin of Error",
                                   -"United States!!Japanese alone!!Margin of Error",
                                   -"United States!!Korean alone!!Margin of Error",
                                   -"United States!!Laotian alone!!Margin of Error",
                                   -"United States!!Pakistani alone!!Margin of Error",
                                   -"United States!!Thai alone!!Margin of Error",
                                   -"United States!!Vietnamese alone!!Margin of Error",
                                   -"United States!!Asian Indian alone or in any combination!!Estimate":-"United States!!Nepalese alone or in any combination!!Margin of Error") |>
  rename("Asian_Indian" = "United States!!Asian Indian alone!!Estimate",
         "Bangladeshi" = "United States!!Bangladeshi alone!!Estimate",
         "Cambodian" = "United States!!Cambodian alone!!Estimate",
         "Chinese" = "United States!!Chinese alone!!Estimate",
         "Chinese_except_Taiwanese" = "United States!!Chinese, except Taiwanese alone!!Estimate",
         "Taiwanese" = "United States!!Taiwanese alone!!Estimate",
         "Filipino" = "United States!!Filipino alone!!Estimate",
         "Hmong" = "United States!!Hmong alone!!Estimate",
         "Indonesian" = "United States!!Indonesian alone!!Estimate",
         "Japanese" = "United States!!Japanese alone!!Estimate",
         "Korean" = "United States!!Korean alone!!Estimate",
         "Laotian" = "United States!!Laotian alone!!Estimate",
         "Pakistani" = "United States!!Pakistani alone!!Estimate",
         "Thai" = "United States!!Thai alone!!Estimate",
         "Vietnamese" = "United States!!Vietnamese alone!!Estimate")

educational_attainment <- acs2022select[c(97:103, 106), ] |> rename("educational_attainment" = "Label (Grouping)")

educational_attainment

educational_attainment <- acs2022select[c(98:103, 106), ] |> rename("educational_attainment" = "Label (Grouping)")

educational_attainment

educational_attainment$Asian_Indian <- c(207475, 220050, 264060, 1002799, 1452330, 2936097, 2451986)

educational_attainment$Chinese <- c(496717, 453086, 419524, 919598, 1067271, 2859481, 1986869)

educational_attainment$Hmong <- c(30487, 46735, 55498, 37424, 12414, 152072, 49838)
```

```{r}
Chinese_plot <- educational_attainment |>
  ggplot(aes(x = Chinese,
             y = educational_attainment)) +
  labs(x = "Chinese population",
       y = "Level of educational attainment",
       title = "Educational Attainment of Chinese Americans") +
  geom_col()

print(Chinese_plot)

Hmong_plot <- educational_attainment |>
  ggplot(aes(x = Hmong,
             y = educational_attainment)) +
  labs(x = "Hmong population",
       y = "Level of educational attainment",
       title = "Educational Attainment of Hmong Americans") +
  geom_col()

print(Hmong_plot)
```


```{r}
educational_attainment_percentages <- acs2022select[c(97:103, 106), ] |> rename("educational_attainment" = "Label (Grouping)")
educational_attainment_percentages |> knitr::kable()
```

From these first two visualizations, it is clear that the first part of my hypothesis — that Chinese Americans and Hmong Americans will have quite disparate levels of educational attainment — is likely true. The data point which exemplifies this best is the percentage of these respective populations which hold a bachelor's degree or higher. Whereas 59.2% of the Chinese American population holds a bachelor's degree or higher, about half of that percentage, or 27.3%, of the Hmong American population holds a bachelor's degree or higher. This large disparity indicates that the average/aggregate of the two will likely fall somewhere near the percentage of the general American population that holds a bachelor's degree or higher (35.7%, according to the 2020 U.S. Census).

## Regression

```{r}
regression_data = data.frame(asian = c(0.592, 0.273),
                             general = 0.357)
model = lm(asian ~ general, regression_data)

model
```

## Conclusion

In this regression, the dependent variable is the "aggregate" percentage of the AANHPI population that holds a bachelor's degree or higher (as determined by taking the average of the respective percentages of the Chinese American and Hmong American population that hold a bachelor's degree or higher). The independent variable is the percentage of the total U.S. population that holds a bachelor's degree or higher. The resulting coefficient is 0.4325, which is statistically significant but cannot be interpreted causally as this was a cross-sectional experiment designed to visualize the disparities between various AANHPI ethnicities' levels of educational attainment and then compare this against that of the total U.S. population, not to attribute any sort of causation between these various percentages.

The coefficient of 0.4325 is a positive, relatively insignificant coefficient,  which I hypothesized the results would return. This coefficient proves that the aggregate AANHPI level of educational attainment is somewhere in between that of the average citizen, effectively proving that AANHPI population is not a monolith (i.e. fares much better or much worse than the general U.S. population), at least when it comes to educational attainment.
